cardinalR: Generating interesting high-dimensional data structures

### 2025-10-13 paper-cardinalR.Rmd
Added initial NLDR summary statistics

@@

# Summarize key performance metrics
summary_clust <- tibble::tibble(
  Metric = c("k-means", "Hierarchical", "Model-based"),
  `wb.ratio` = round(c(comp_kmeans$wb.ratio, comp_hclust$wb.ratio, comp_mclust$wb.ratio), 2),
  `Dunn Index` = round(c(comp_kmeans$dunn, comp_hclust$dunn, comp_mclust$dunn), 2),
  `Corrected rand` = round(c(comp_kmeans$corrected.rand, comp_hclust$corrected.rand, comp_mclust$corrected.rand), 2),
  VI = round(c(comp_kmeans$vi, comp_hclust$vi, comp_mclust$vi), 2)
)
```

>>>>

```{r}
summary_dr <- read_rds("data/five_clusts/dr_summary.rds")

summary_dr <- summary_dr |>
  mutate(across(where(is.numeric), ~ round(.x, 2)))
```

```{r summarydr-tb-html, eval=knitr::is_html_output()}
summary_dr |> 
  kable(caption = "Comparison of clustering performance metrics (within–between ratio (wb.ratio), Dunn index, Corrected Rand index, and variation of information (VI) across $k$-means, hierarchical, and model-based clustering methods.", col.names = c("Method", "svm", "cte", "rte", "global_score"), tab.pos = "H", linesep = "") 
```

```{r summarydr-tb-pdf, eval=knitr::is_latex_output()}
summary_dr |> 
  kable(caption = "Comparison of clustering performance metrics (within–between ratio (wb.ratio), Dunn index, Corrected Rand index, and variation of information (VI) across $k$-means, hierarchical, and model-based clustering methods.", format="latex", col.names = c("Method", "svm", "cte", "rte", "global_score"), booktabs = T, table.pos = "H", linesep = "")  |>
  column_spec(1, width = "2cm") |>
  column_spec(2, width = "2cm") |>
  column_spec(3, width = "2cm") |>
  column_spec(4, width = "2cm") |>
  column_spec(5, width = "2cm")
```

@@

# Summarize key performance metrics
summary_clust <- tibble::tibble(
  Metric = c("k-means", "Hierarchical", "Model-based"),
  `wb.ratio` = c(comp_kmeans$wb.ratio, comp_hclust$wb.ratio, comp_mclust$wb.ratio),
  `Dunn Index` = c(comp_kmeans$dunn, comp_hclust$dunn, comp_mclust$dunn),
  `Corrected rand` = c(comp_kmeans$corrected.rand, comp_hclust$corrected.rand, comp_mclust$corrected.rand),
  VI = c(comp_kmeans$vi, comp_hclust$vi, comp_mclust$vi)
)

summary_clust <- summary_clust |>
  mutate(across(where(is.numeric), ~ round(.x, 2)))
```

### 2025-10-13 paper-cardinalR.tex
Added initial NLDR summary statistics

>>>>

\caption{\label{tab:summarydr-tb-pdf}Comparison of clustering performance metrics (within–between ratio (wb.ratio), Dunn index, Corrected Rand index, and variation of information (VI) across $k$-means, hierarchical, and model-based clustering methods.}
\centering
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2cm}>{\raggedleft\arraybackslash}p{2cm}>{\raggedleft\arraybackslash}p{2cm}>{\raggedleft\arraybackslash}p{2cm}>{\raggedleft\arraybackslash}p{2cm}}
\toprule
Method & svm & cte & rte & global\_score\\
\midrule
tSNE & 0.98 & 0.42 & 0.70 & 0.46\\
UMAP & 0.98 & 0.62 & 0.73 & 0.58\\
PHATE & 0.90 & 0.47 & 0.72 & 0.55\\
TriMAP & 0.93 & 0.58 & 0.75 & 0.59\\
PaCMAP & 0.99 & 0.24 & 0.72 & 0.54\\
PCA & 0.78 & 0.53 & 0.83 & 1.00\\
\bottomrule
\end{tabular}
\end{table}
